\textbf{ Spectrophotometric parallaxes with linear models:\\ Accurate distances for luminous red-giant stars }

### 2018-10-21 paper/sp.tex
arxiver, cmd, refs

% To-do items
% -----------
% - check with WISE experts that we have the correct WISE photometry
% - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?
% - are we clear that the map is just a demo and just shows the 10-percent-ish region of the MW disk?

@@

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{TBD}
%\revised{TBD}
%\accepted{TBD}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
%\submitjournal{AAS Journals}

\shorttitle{spectrophotometric parallax}
\shortauthors{hogg, eilers, rix}
%\watermark{kittens}

@@

\newpage
\begin{longrotatetable}
\startlongtable
\begin{deluxetable*}{ccccccc}
\tablecaption{HOGG and ACE TBD: The generated spectrophotometric parallaxes.
This table is just a snippet; the full table is available online (HOGG HOW?)\label{tab:output}}
\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
\startdata
2M00000002+7417074 & 0.3191 & 0.0329 & 0.3185 & 0.0164 & 0 & B \\
2M00000317+5821383 & 0.3643 & 0.0510 & 0.4428 & 0.0183 & 0 & A \\
2M00000546+6152107 & 0.3154 & 0.0351 & 0.3292 & 0.0164 & 0 & B \\
2M00000797+6436119 & 0.0771 & 0.0841 & 0.1943 & 0.0194 & 1 & B \\
2M00001719+6221324 & 0.1395 & 0.0439 & 0.1564 & 0.0128 & 0 & B \\
2M00002005+5703467 & 0.3003 & 0.0247 & 0.2372 & 0.0203 & 0 & B \\
2M00002036+6153416 & 0.0657 & 0.0637 & 0.1347 & 0.0131 & 1 & A \\
2M00002118+6136420 & 0.1936 & 0.0323 & 0.2173 & 0.0133 & 1 & B \\
2M00002227+6223341 & 0.1965 & 0.0450 & 0.1461 & 0.0089 & 0 & B \\
2M00002472+5518473 & 0.0947 & 0.0207 & 0.1473 & 0.0096 & 1 & B \\
2M00002908+6140446 & 0.1675 & 0.0460 & 0.1914 & 0.0176 & 1 & A \\
2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.0290 & 1 & A \\
... & ... & ... & ... & ... & ... & ...  \\
 \enddata
\tablecomments{Columns show...}
\end{deluxetable*}
\end{longrotatetable}
\newpage

@@

HOGG: Discussion here of CMD and AGB stars.
\begin{figure}
\centering
~\hfill\includegraphics[width=0.8\textwidth]{./CMD4.pdf}\hfill~
\caption{HOGG color-magnitude-diagram.\label{fig:cmd}}
\end{figure}

This \documentname\ and project delivers greatly improved distance estimates for
luminous red-giant stars.
The purpose of all this is to make global maps of the Milky Way.
Just as a demonstration, we show kinematics and element abundances for stars as a function
of position in the Milky Way disk in \figurename~\ref{fig:disk}.
In this visualization, it is possible to see the rotation of the disk and the
kinematic center of the Galaxy; that might even provide a distance estimate to the Galactic
Center.
It also shows extremely strong chemical gradients.
\begin{figure}
\includegraphics[width=\textwidth]{map.pdf}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  Each arrow represents the three-space Galactocentric velocity of each star
  in the sample, projected onto the Galactic disk plane.
  To make these velocities, raw \gaia\ proper motions and \apogee\ radial velocities
  were used (in a standard Galactocentric transformation), so they are
  not in any way noise-deconvolved.
  Each arrow has a color corresponding to the \apogee\ reported metallicity.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. Interpretation of
  the quantitative information in this plot requires care.\label{fig:disk}}
\end{figure}
In a companion paper (\citealt{eilers}), we are using this information to estimate
the (asymmetric-drift-corrected) circular-velocity curve for the Milky Way.

@@

\begin{thebibliography}{}
\bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
\bibitem[Allende Prieto \etal(2008)]{aapogee} Allende Prieto, C., Majewski, S.~R., Schiavon, R., \etal\ 2008, Astronomische Nachrichten, 329, 1018 
\bibitem[Arenou \etal(2018)]{arenou} Arenou, F., Luri, X., Babusiaux, C., et al.\ 2018, \aap, 616, A17 
\bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
\bibitem[Astropy Collaboration \etal(2018)]{astropy2} The Astropy Collaboration, Price-Whelan, A.~M., Sip{\H o}cz, B.~M., \etal\ 2018, arXiv:1801.02634 
\bibitem[Bailer-Jones \etal(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
\bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
\bibitem[Byrd \etal(1995)]{lbfgsb} Byrd, R.~H., Lu,~P., Nocedal,~J., \& Zhu,~C. 1995, SIAM J. Sci. Comput., 16, 1190
\bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
\bibitem[Eilers \etal(2018)]{eilers} Eilers,~A.~C., Hogg, D.~W., Rix, H.-W., \& Ness, M. 2018, in preparation
\bibitem[Gaia Collaboration \etal(2016)]{gaia} Gaia Collaboration, Prusti, T., de Bruijne, J.~H.~J., \etal\ 2016, \aap, 595, A1 
\bibitem[Gaia Collaboration \etal(2018)]{gaiadr2} Gaia Collaboration, Brown, A.~G.~A., Vallenari, A., \etal\ 2018, arXiv:1804.09365 
\bibitem[Garc{\'{\i}}a P{\'e}rez \etal(2016)]{aspcap} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., \etal\ 2016, \aj, 151, 144 
\bibitem[Ho \etal(2017)]{ho} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., \etal\ 2017, \apj, 836, 5 
\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W.\ 2018, arXiv:1804.07766
\bibitem[Hunter(2007)]{matplotlib} Hunter, J.~D.\ 2007, CISE, 9(3), 90
\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001, \url{http://www.scipy.org/}
\bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 
\bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
\bibitem[Majewski \etal(2017)]{apogee} Majewski, S.~R., Schiavon, R.~P., Frinchaboy, P.~M., \etal\ 2017, \aj, 154, 94 
\bibitem[Marrese \etal(2018)]{xmatch} Marrase, P.~M. \etal\ 2018, in preparation
\bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
\bibitem[Mora \etal(2018)]{gaiaarchive} Mora, A., Gonz{\'a}lez-N{\'u}{\~n}ez, J., Baines, D., \etal\ 2018, IAU Symposium, 330, 35 
\bibitem[Ness \etal(2015)]{cannon} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16 
\bibitem[Ness \etal(2016)]{nessage} Ness, M., Hogg, D.~W., Rix, H.-W., \etal\ 2016, \apj, 823, 114 
\bibitem[Ness \etal(2018)]{nessdopp} Ness, M., Rix, H.-W., Hogg, D.~W., \etal\ 2018, \apj, 853, 198 
\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing
\bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29
\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
\bibitem[Wilson \etal(2010)]{wapogee} Wilson, J.~C., Hearty, F., Skrutskie, M.~F., \etal\ 2010, \procspie, 7735, 77351C 
\bibitem[Wright \etal(2010)]{wise} Wright, E.~L., Eisenhardt, P.~R.~M., Mainzer, A.~K., \etal\ 2010, \aj, 140, 1868 
\bibitem[Zinn \etal(2018)]{zinn} Zinn, J.~C., Pinsonneault, M.~H., Huber, D., \& Stello, D.\ 2018, arXiv:1805.02650 
\end{thebibliography}

>>>>

%@arxiver{residuals.pdf,map.pdf}

% To-do items
% -----------
% - check with WISE experts that we have the correct WISE photometry
% - ACE: What fraction of spectral pixels are (typically, or in each training set) zeroed out?

@@

\shorttitle{spectrophotometric parallax}
\shortauthors{hogg, eilers, rix}

@@

\begin{longrotatetable}
\startlongtable
\begin{deluxetable*}{ccccccc}
\tablecaption{The generated spectrophotometric parallaxes and their uncertainties..
This \tablename\ is just a snippet; the full catalog of all 44784 objects in the combined Parent Samples
is available online. In order to reproduce the {\figurename s} in this \documentname, this catalog
should be joined to the \gaia, \zmass, \wise\, and \apogee\ Catalogs by \zmass\ ID.\label{tab:output}}
\tablehead{\dcolhead{\texttt{2MASS\_ID}} & \dcolhead{\texttt{Gaia\_parallax}} & \dcolhead{\texttt{Gaia\_parallax\_err}} & \dcolhead{\texttt{spec\_parallax}} & \dcolhead{\texttt{spec\_parallax\_err}} & \dcolhead{\texttt{training\_set}} & \dcolhead{\texttt{sample}}}
\startdata
2M00000002+7417074 & 0.3191 & 0.0329 & 0.3185 & 0.0164 & 0 & B \\
2M00000317+5821383 & 0.3643 & 0.0510 & 0.4428 & 0.0183 & 0 & A \\
2M00000546+6152107 & 0.3154 & 0.0351 & 0.3292 & 0.0164 & 0 & B \\
2M00000797+6436119 & 0.0771 & 0.0841 & 0.1943 & 0.0194 & 1 & B \\
2M00001719+6221324 & 0.1395 & 0.0439 & 0.1564 & 0.0128 & 0 & B \\
2M00002005+5703467 & 0.3003 & 0.0247 & 0.2372 & 0.0203 & 0 & B \\
2M00002036+6153416 & 0.0657 & 0.0637 & 0.1347 & 0.0131 & 1 & A \\
2M00002118+6136420 & 0.1936 & 0.0323 & 0.2173 & 0.0133 & 1 & B \\
2M00002227+6223341 & 0.1965 & 0.0450 & 0.1461 & 0.0089 & 0 & B \\
2M00002472+5518473 & 0.0947 & 0.0207 & 0.1473 & 0.0096 & 1 & B \\
2M00002908+6140446 & 0.1675 & 0.0460 & 0.1914 & 0.0176 & 1 & A \\
2M00003061+5820590 & 0.1669 & 0.0339 & 0.3502 & 0.0290 & 1 & A \\
... & ... & ... & ... & ... & ... & ...  \\
 \enddata
\tablecomments{Columns show...}
\end{deluxetable*}
\end{longrotatetable}
\newpage

@@

This \documentname\ and project delivers greatly improved distance estimates for
luminous red-giant stars, more luminous than the red clump.
The purpose of all this is to make global maps of the Milky Way.
Just as a demonstration, we show kinematics and element abundances for stars as a function
of position in the Milky Way disk in \figurename~\ref{fig:disk}.
In this visualization, it is possible to see the rotation of the disk and the
kinematic center of the Galaxy; that might even provide a distance estimate to the Galactic
Center.
It also shows extremely strong chemical gradients.
\begin{figure}
\includegraphics[width=\textwidth]{map.pdf}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  Each arrow represents the three-space Galactocentric velocity of each star
  in the sample, projected onto the Galactic disk plane.
  To make these velocities, raw \gaia\ proper motions and \apogee\ radial velocities
  were used (in a standard Galactocentric transformation), so they are
  not in any way noise-deconvolved.
  Each arrow has a color corresponding to the \apogee\ reported metallicity.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. The stars shown in this
  map have distance uncertainties on the order of ten percent, so at significant
  heliocentric distances, the interpretation of
  the quantitative information in this plot requires care.\label{fig:disk}}
\end{figure}
In a companion paper (\citealt{eilers}), we are using this information to estimate
the (asymmetric-drift-corrected) circular-velocity curve for the Milky Way.

@@

\begin{thebibliography}{}
\bibitem[Abolfathi \etal(2018)]{dr14} Abolfathi, B., Aguado, D.~S., Aguilar, G., \etal\ 2018, \apjs, 235, 42 
\bibitem[Allende Prieto \etal(2008)]{aapogee} Allende Prieto, C., Majewski, S.~R., Schiavon, R., \etal\ 2008, Astronomische Nachrichten, 329, 1018 
\bibitem[Arenou \etal(2018)]{arenou} Arenou, F., Luri, X., Babusiaux, C., \etal\ 2018, \aap, 616, A17 
\bibitem[Astropy Collaboration \etal(2013)]{astropy} Astropy Collaboration, Robitaille, T.~P., Tollerud, E.~J., \etal\ 2013, \aap, 558, A33 
\bibitem[Astropy Collaboration \etal(2018)]{astropy2} The Astropy Collaboration, Price-Whelan, A.~M., Sip{\H o}cz, B.~M., \etal\ 2018, arXiv:1801.02634 
\bibitem[Bailer-Jones \etal(2018)]{calj} Bailer-Jones, C.~A.~L., Rybizki, J., Fouesneau, M., Mantelet, G., \& Andrae, R.\ 2018, \aj, 156, 58 
\bibitem[Blanton \etal(2017)]{sdssiv} Blanton, M.~R., Bershady, M.~A., Abolfathi, B., \etal\ 2017, \aj, 154, 28 
\bibitem[Byrd \etal(1995)]{lbfgsb} Byrd, R.~H., Lu,~P., Nocedal,~J., \& Zhu,~C. 1995, SIAM J. Sci. Comput., 16, 1190
\bibitem[Casey \etal(2016)]{casey} Casey, A.~R., Hogg, D.~W., Ness, M., \etal\ 2016, arXiv:1603.03040 
\bibitem[Eilers \etal(2018)]{eilers} Eilers,~A.~C., Hogg, D.~W., Rix, H.-W., \& Ness, M. 2018, in preparation
\bibitem[Gaia Collaboration \etal(2016)]{gaia} Gaia Collaboration, Prusti, T., de Bruijne, J.~H.~J., \etal\ 2016, \aap, 595, A1 
\bibitem[Gaia Collaboration \etal(2018)]{gaiadr2} Gaia Collaboration, Brown, A.~G.~A., Vallenari, A., \etal\ 2018, arXiv:1804.09365 
\bibitem[Garc{\'{\i}}a P{\'e}rez \etal(2016)]{aspcap} Garc{\'{\i}}a P{\'e}rez, A.~E., Allende Prieto, C., Holtzman, J.~A., \etal\ 2016, \aj, 151, 144 
\bibitem[Helmi \etal(2018)]{helmi} Helmi, A., Babusiaux, C., Koppelman, H.~H., \etal\ 2018, arXiv:1806.06038 
\bibitem[Ho \etal(2017)]{ho} Ho, A.~Y.~Q., Ness, M.~K., Hogg, D.~W., \etal\ 2017, \apj, 836, 5 
\bibitem[Hogg(2018)]{gaialf} Hogg, D.~W.\ 2018, arXiv:1804.07766
\bibitem[Hunter(2007)]{matplotlib} Hunter, J.~D.\ 2007, CISE, 9(3), 90
\bibitem[Jones \etal(2001)]{scipy} Jones, E., Oliphant, E., Peterson, P. \etal\ 2001, \url{http://www.scipy.org/}
\bibitem[Kollmeier \etal(2017)]{sdssv} Kollmeier, J.~A., Zasowski, G., Rix, H.-W., \etal\ 2017, arXiv:1711.03234 
\bibitem[Lindegren \etal(2018)]{lindegren} Lindegren, L., Hernandez, J., Bombrun, A., \etal\ 2018, arXiv:1804.09366 
\bibitem[Lutz \& Kelker(1973)]{lk} Lutz, T.~E., \& Kelker, D.~H.\ 1973, \pasp, 85, 573 
\bibitem[Majewski \etal(2017)]{apogee} Majewski, S.~R., Schiavon, R.~P., Frinchaboy, P.~M., \etal\ 2017, \aj, 154, 94 
\bibitem[Marrese \etal(2018)]{xmatch} Marrase, P.~M. \etal\ 2018, in preparation
\bibitem[Martig \etal(2016)]{martig} Martig, M., Fouesneau, M., Rix, H.-W., \etal\ 2016, \mnras, 456, 3655 
\bibitem[Mora \etal(2018)]{gaiaarchive} Mora, A., Gonz{\'a}lez-N{\'u}{\~n}ez, J., Baines, D., \etal\ 2018, IAU Symposium, 330, 35 
\bibitem[Ness \etal(2015)]{cannon} Ness, M., Hogg, D.~W., Rix, H.-W., Ho, A.~Y.~Q., \& Zasowski, G.\ 2015, \apj, 808, 16 
\bibitem[Ness \etal(2016)]{nessage} Ness, M., Hogg, D.~W., Rix, H.-W., \etal\ 2016, \apj, 823, 114 
\bibitem[Ness \etal(2018)]{nessdopp} Ness, M., Rix, H.-W., Hogg, D.~W., \etal\ 2018, \apj, 853, 198 
\bibitem[Oliphant(2006)]{numpy} Oliphant, T.~E.\ 2006, \emph{A guide to NumPy}, Trelgol Publishing
\bibitem[P\'erez \& Granger(2007)]{ipython} P\'erez, F. \& Granger, B.~E.\ 2007, CISE, 9(3), 29
\bibitem[Skrutskie \etal(2006)]{zmass} Skrutskie, M.~F., Cutri, R.~M., Stiening, R., \etal\ 2006, \aj, 131, 1163 
\bibitem[Trick \etal(2018)]{trick} Trick, W.~H., Coronado, J., \& Rix, H.-W.\ 2018, arXiv:1805.03653 
\bibitem[Wilson \etal(2010)]{wapogee} Wilson, J.~C., Hearty, F., Skrutskie, M.~F., \etal\ 2010, \procspie, 7735, 77351C 
\bibitem[Wright \etal(2010)]{wise} Wright, E.~L., Eisenhardt, P.~R.~M., Mainzer, A.~K., \etal\ 2010, \aj, 140, 1868 
\bibitem[Zinn \etal(2018)]{zinn} Zinn, J.~C., Pinsonneault, M.~H., Huber, D., \& Stello, D.\ 2018, arXiv:1805.02650 
\end{thebibliography}
